---
title: "Analysis_main"
author: "TR"
date: "`r Sys.Date()`"
output: html_document
---
**This R markdown file contains all relevant analysis steps of the manuscript titled "Ecotype diversification as mechanism for coexistence in extreme environments – the case of northern pike (Esox lucius) in brackish lagoons", including data preparation of elemental transect data, genetics data and growth data (growth data was produced via OtoJ macro of imageJ´s objectJ plugin). Not included are the assignment of element transect data to individual years, raw data preparation of SIMS and LA-ICPMS analysis and plotting of individual-level transects for evaluation of data quality, outliers or vaterite inclusions (indicated by Sr - Mg interaction)**

**All chunks except for the packages chunk have been set to eval = FALSE, all chunks can be run individually**

**Note that this file has not been optimized for any kind of output, if you wish to produce an html or word document, you may need to adapt figure and table outputs manually**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(ggridges)
library(lme4)
library(nlme)
library(lmerTest)
library(emmeans)
library(MuMIn)
library(visreg)
library(car)
library(sjPlot)
library(knitr)
library(car)
library(ggpubr)
library(mclust)
library(dtwclust)
library(vegan)
library(mvabund)
library(RVAideMemoire)
library(MASS)
library(flextable)
library(gridGraphics)
library(lubridate)
library(factoextra)
library(data.table)
library(gridExtra)
library(loo)
library(bayesplot)
library("rstan")
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

# Life history phenotypes
## Data prep for niche analysis
### Notes on preparation of Sr and d18O data
Five fish excluded: 
"BH-01701" -> Misplaced transect,
"BH-01741" -> Vaterite,
"BH-01908" -> Vaterite,
"BH-90962" -> Lost during SIMS preprocessing,
"BH-93190" -> no SIMS transect (lightning strike occurred during analysis)
```{r marry d18O data to Sr data, include=FALSE, eval=FALSE}
rm(list = ls())

#Read in LA-ICPMS and SIMS data
LA <- fread("Data/pike/elements/Pike-LA-ICPMS-years_whole-sample_with-TL.txt")
d18O <- fread("Data/pike/elements/Pike-d18O-years_whole-sample_with-TL.txt")

#Clean up data & assign capture locations to larger study areas
#Clean up Trace element data
LA_clean <- LA%>%
  filter(LA_dist>=50)%>%
  filter(comment=="OK")%>%
  transmute(fish_ID = fishID,
            age=age,
            capt = area,
            area = case_when(area %in% "Peene" ~ "Freshwater",
                             area %in% "Barthe" ~ "Freshwater",
                             area %in% "NHG" ~ "Freshwater",
                             area %in% "Sehrowbach" ~ "Freshwater",
                             area %in% "Bandycksgraben" ~ "Freshwater",
                             area %in% "Ziese" ~ "Freshwater",
                             area %in% "KB" ~ "WRB",
                             area %in% "SB" ~ "WRB",
                             area %in% "WB" ~ "NRB",
                             area %in% "BEG" ~ "NRB", 
                             area %in% "KJB" ~ "NRB",
                             area %in% "GJB" ~ "NRB",
                             area %in% "GB" ~ "GB",
                             area %in% "KD" ~ "control",
                             TRUE ~ "NA"),
            Sr = Sr,
            Na = Na,
            Mg = Mg,
            Ba = Ba,
            Mn = Mn,
            LA_dist = LA_dist,
            lifeyear=lifeyear)

#Clean up d18O data
d18O_clean <- d18O%>%
  filter(comment=="OK")%>%
  transmute(fish_ID = fishID,
            age=age,
            capt = area,
            area = case_when(area %in% "Peene" ~ "Freshwater",
                             area %in% "Barthe" ~ "Freshwater",
                             area %in% "NHG" ~ "Freshwater",
                             area %in% "Sehrowbach" ~ "Freshwater",
                             area %in% "Badendycksgraben" ~ "Freshwater",
                             area %in% "Ziese" ~ "Freshwater",
                             area %in% "KB" ~ "WRB",
                             area %in% "SB" ~ "WRB",
                             area %in% "WB" ~ "NRB",
                             area %in% "BEG" ~ "NRB", 
                             area %in% "KJB" ~ "NRB",
                             area %in% "GJB" ~ "NRB",
                             area %in% "GB" ~ "GB",
                             area %in% "KD" ~ "Control",
                             TRUE ~ "NA"),
            d18O = d18O,
            sims_dist = sims_dist,
            lifeyear=lifeyear)

LA_clean <- LA_clean %>% drop_na()

#Smooth trace element data to remove noise
LA_smooth <- LA_clean%>%group_by(fish_ID)%>%
  transmute(fish_ID=fish_ID,
            age=age,
            capt=capt,
            area=area,
            Sr=as.numeric(zoo::rollmean(Sr, k=9, fill=NA, align = "left")),
            Na=as.numeric(zoo::rollmean(Na, k=9, fill=NA, align = "left")),
            Mg=as.numeric(zoo::rollmean(Mg, k=9, fill=NA, align = "left")),
            Ba=as.numeric(zoo::rollmean(Ba, k=9, fill=NA, align = "left")),
            Mn=as.numeric(zoo::rollmean(Mn, k=9, fill=NA, align = "left")),
            LA_dist=as.numeric(LA_dist),
            lifeyear=lifeyear)

#Order data sets for joining
d18O_clean <- d18O_clean[order(d18O_clean$fish_ID),]
LA_smooth <- LA_smooth[order(LA_smooth$fish_ID),]
LA_smooth <- data.frame(LA_smooth)

#Convert to data frame, so approx (), as approx() behaves weird on data tables
d18O_clean <- as.data.frame(d18O_clean)
LA_clean <- as.data.frame(LA_clean)

#reinterpolate Sr (higher spatial resolution ~5.5 um) to d18O distances (lower spatial resolution, ~35 um) before joining
IDOto <- unique(d18O_clean$fish_ID)
dfE <- data.frame()
for (i in 1:length(IDOto)) {
  Sr <- data.frame(approx(LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"LA_dist"],
              LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"Sr"],
              n = length(d18O_clean[which(d18O_clean$fish_ID==IDOto[i]),"sims_dist"]),
              rule = 1, method = "linear", ties = "ordered"))
  Na <- data.frame(approx(LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"LA_dist"],
              LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"Na"],
              n = length(d18O_clean[which(d18O_clean$fish_ID==IDOto[i]),"sims_dist"]),
              rule = 1, method = "linear", ties = "ordered"))
  Mg <- data.frame(approx(LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"LA_dist"],
              LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"Mg"],
              n = length(d18O_clean[which(d18O_clean$fish_ID==IDOto[i]),"sims_dist"]),
              rule = 1, method = "linear", ties = "ordered"))
  Ba <- data.frame(approx(LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"LA_dist"],
              LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"Ba"],
              n = length(d18O_clean[which(d18O_clean$fish_ID==IDOto[i]),"sims_dist"]),
              rule = 1, method = "linear", ties = "ordered"))
  Mn <- data.frame(approx(LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"LA_dist"],
              LA_smooth[which(LA_smooth$fish_ID==IDOto[i]),"Mn"],
              n = length(d18O_clean[which(d18O_clean$fish_ID==IDOto[i]),"sims_dist"]),
              rule = 1, method = "linear", ties = "ordered"))
  df <- cbind(Sr=Sr[,2], Na=Na[,2], Mg=Mg[,2], Ba=Ba[,2], Mn=Mn[,2])
  dfE <- rbind(dfE,df)
}

oto.int <- cbind(d18O_clean, dfE)

fwrite(oto.int, file = "Data/pike/elements/Oto_LA-to-d18O_interpolation.csv")
```

```{r covariance analysis d18O vs Sr and data correction, include=FALSE, eval=FALSE}
rm(list = ls())

#read in combined element transect data
oto.int <- fread("Data/pike/elements/Oto_LA-to-d18O_interpolation.csv")

#clean up to remove fish from control lake (but make subset for later join)
oto.int2 <- oto.int%>%
  filter(capt!="KD")%>%
  drop_na()

subset_KD <- subset(oto.int, capt=="KD")

#plot correlation between d18O and Sr
Correlation.Sr <- ggplot(oto.int2,aes(x=Sr*1000, y=d18O))+
  geom_point(size=.5)+
  scale_x_continuous(breaks = seq(0,8,1))+
  scale_y_continuous(breaks = seq(-15,0,1))+
  labs(x="Sr/Ca (mg/g)", 
       y=expression(bold(delta^18*"O"~("VPDB \u2030"))),
       title=expression(bold("Correlation of intraotolith Sr and"~delta^18*"O")))+
  stat_smooth(method = "lm", fullrange = T, size = .8)+
  theme_classic()+
    theme(panel.grid.major = element_line(),
          axis.line = element_line(size = .5, colour = "black", linetype = 1),
          axis.ticks = element_line(size = .5, colour = "black"),
          axis.text = element_text(size = 10, colour = "black"),
          axis.title.x = element_text(size = 10, face = "bold"),
          axis.title.y = element_text(size = 10, face = "bold"),
          legend.text = element_text(size = 5),
          legend.title = element_text(size = 10, face = "bold"),
          plot.title = element_text(size = 15, face = "bold", hjust = .5))

ggsave("Figures/supplement/Correlation_d18O~Sr.png", Correlation.Sr, width = 15, height = 9, units = "cm", dpi = 600, bg="white")

#Test different models
#linear model
lm1 <- lm(d18O~Sr, data = oto.int2)
#exponential decay model
lm2 <- lm(d18O~exp(-Sr), data = oto.int2)

#Test models
anova(lm1, lm2) #exponential decay is not significantly better than a linear model

#summarize final model
summary(lm1)

#Extract residuals
oto.int2$res <- lm1$residuals

#fill in scaled values for d18O in pike from control lake
subset_KD$res <- scale(subset_KD$d18O)
subset_KD <- subset_KD%>%drop_na()

oto.int3 <- rbind(oto.int2, subset_KD)

#check for pattern retention
plot(oto.int3[oto.int3$fish_ID=="BH-01906",]$sims_dist, oto.int3[oto.int3$fish_ID=="BH-01906",]$res, type="l")

fwrite(oto.int3, "Data/pike/elements/Oto_LA-to-d18O_interpolation_Sr-corrected.csv")
```

```{r average corrected d18O over years, include=FALSE, eval=FALSE}
rm(list = ls())

#read in data
oto.int <- fread("Data/pike/elements/Oto_LA-to-d18O_interpolation_Sr-corrected.csv")

#Make lifeyear a factor
oto.int$lifeyear <- as.factor(oto.int$lifeyear)

#calculate average number of d18O points per year
yearmean <- oto.int%>%group_by(fish_ID,lifeyear)%>%summarise(n=n())
mean(yearmean$n) #mean of 7.8 --> smoothing data with 7 points rolling mean

oto.int2 <- oto.int%>%
  group_by(fish_ID)%>%
  transmute(fish_ID=fish_ID,
            age=age,
            capt=capt,
            area=area,
            d18O=zoo::rollmean(d18O, k=7, fill = NA, align = "left"),
            res=zoo::rollmean(res, k=7, fill = NA, align = "left"),
            Sr=Sr,
            Na=Na,
            Mg=Mg,
            Ba=Ba,
            Mn=Mn,
            year=lifeyear,
            dist=sims_dist)

#check whether smoothing averaged over seasonal d18O signal
par(mfrow=c(1,2))
plot(oto.int[oto.int$fish_ID=="BH-01855",]$sims_dist, oto.int[oto.int$fish_ID=="BH-01855",]$d18O, type="l")
plot(oto.int2[oto.int2$fish_ID=="BH-01855",]$dist, oto.int2[oto.int2$fish_ID=="BH-01855",]$d18O, type="l")

#Save data
fwrite(oto.int2, "Data/pike/elements/Oto_LA-to-d18O_interpolation_Sr-corrected_smooth.csv")
```
## Lifelong niche (individual and population level)
```{r calculate profile plots for individuals, include=FALSE, eval=FALSE}
rm(list = ls())

#read in raw data
LA <- fread("Data/pike/elements/Pike-LA-ICPMS-years_whole-sample_with-TL.txt")
d18O <- fread("Data/pike/elements/Pike-d18O-years_whole-sample_with-TL.txt")

#Convert to data frame
LA <- as.data.frame(LA)
d18O <- as.data.frame(d18O)

LA <- LA%>%
  transmute(ID = ID,
            capt = area,
            area = case_when(area %in% "Peene" ~ "Freshwater",
                             area %in% "Barthe" ~ "Freshwater",
                             area %in% "NHG" ~ "Freshwater",
                             area %in% "Sehrowbach" ~ "Freshwater",
                             area %in% "Badendycksgraben" ~ "Freshwater",
                             area %in% "Ziese" ~ "Freshwater",
                             area %in% "KB" ~ "Marine",
                             area %in% "SB" ~ "Marine",
                             area %in% "WB" ~ "Marine",
                             area %in% "BEG" ~ "Marine", 
                             area %in% "KJB" ~ "Marine",
                             area %in% "GJB" ~ "Marine",
                             area %in% "GB" ~ "Marine",
                             area %in% "KD" ~ "control",
                             TRUE ~ "NA"),
            Sr = Sr,
            Na = Na,
            Mg = Mg,
            Ba = Ba,
            Mn = Mn,
            LA_dist = LA_dist)

d18O <- d18O%>%
  transmute(ID = ID,
            capt = area,
            area = case_when(area %in% "Peene" ~ "Freshwater",
                             area %in% "Barthe" ~ "Freshwater",
                             area %in% "NHG" ~ "Freshwater",
                             area %in% "Sehrowbach" ~ "Freshwater",
                             area %in% "Bandycksgraben" ~ "Freshwater",
                             area %in% "Ziese" ~ "Freshwater",
                             area %in% "KB" ~ "Marine",
                             area %in% "SB" ~ "Marine",
                             area %in% "WB" ~ "Marine",
                             area %in% "BEG" ~ "Marine", 
                             area %in% "KJB" ~ "Marine",
                             area %in% "GJB" ~ "Marine",
                             area %in% "GB" ~ "Marine",
                             area %in% "KD" ~ "Freshwater",
                             TRUE ~ "NA"),
            d18O = d18O,
            sims_dist = sims_dist)


LA <- LA %>% drop_na()
d18O <- d18O%>%drop_na()
# Loop--------------------------------------------------------------------------

mypallete <- viridis::viridis(4, direction = 1)

IDOto <- unique(d18O$ID)

for (i in 1:length(IDOto)) {
  png(paste0("Figures/supplement/individual plots/", IDOto[i], ".png"), width = 1200)
  par(mfrow=c(1,2))
  plot(LA[which(LA$ID == IDOto[i]), "LA_dist",], 
       LA[which(LA$ID == IDOto[i]), "Sr",], main = "", 
       xlab =  "", ylab = "", 
       xlim = c(0, max(LA[which(LA$ID == IDOto[i]), "LA_dist",], na.rm = T)),
       ylim = c(0, max(LA$Sr, na.rm = T)),
       col.axis = "black", type = "l", col = "orange", lwd = 4, 
       fg = "black", axes = FALSE, pch = 19)
  axis(1, pos = mean(LA[LA$area == "Freshwater",]$Sr, na.rm = T), 
       at = c(0, max(LA[which(LA$ID == IDOto[i]), "LA_dist",], na.rm = T)-20,
              max(LA$LA_dist)),
       labels = c("", "Freshwater", ""), lwd.ticks = 0, tck = 0, 
       col = mypallete[3], lty = 1, lwd = 3, mgp = c(0,0.5,0), 
       cex.axis=2, font = 2)
  axis(1, pos = mean(LA[LA$area == "Freshwater",]$Sr, na.rm = T)+
         sd(LA[LA$area == "Freshwater",]$Sr, na.rm = T), 
       at = c(0, 1200, max(LA$LA_dist)),
       labels = c("", "", ""), lwd.ticks = 0, tck = 0, 
       col = mypallete[3], lty = 3, lwd = 3, mgp = c(0,0,0))
  axis(1, pos = mean(LA[LA$area == "Freshwater",]$Sr, na.rm = T)-
         sd(LA[LA$area == "Freshwater",]$Sr, na.rm = T), 
       at = c(0, 1200, max(LA$LA_dist)),
       labels = c("", "", ""), lwd.ticks = 0, tck = 0, 
       col = mypallete[3], lty = 3, lwd = 3, mgp = c(0,0,0))
  axis(1, pos = mean(LA[LA$area == "Marine",]$Sr, na.rm = T), 
       at = c(0, max(LA[which(LA$ID == IDOto[i]), "LA_dist",], na.rm = T)-20,
                   max(LA$LA_dist)),
       labels = c("", "Brackish", ""), lwd.ticks = 0, tck = 0, 
       col = mypallete[1], lty = 1, lwd = 3, mgp = c(0,0.5,0), 
       cex.axis=2, font = 2)
  axis(1, pos = mean(LA[LA$area == "Marine",]$Sr, na.rm = T)-
         sd(LA[LA$area == "Marine",]$Sr, na.rm = T), 
       at = c(0, 1200, max(LA$LA_dist)),
       labels = c("", "", ""), lwd.ticks = 0, tck = 0, 
       col = mypallete[1], lty = 3, lwd = 3, mgp = c(0,0,0))
  axis(1, pos = mean(LA[LA$area == "Marine",]$Sr, na.rm = T)+
         sd(LA[LA$area == "Marine",]$Sr, na.rm = T), 
       at = c(0, 1200, max(LA$LA_dist)),
       labels = c("", "", ""), lwd.ticks = 0, tck = 0, 
       col = mypallete[1], lty = 3, lwd = 3, mgp = c(0,0,0))
  axis(1, pos = 0, at = seq(0, max(LA$LA_dist), 500),
       labels = seq(0, max(LA$LA_dist), 500), lwd.ticks = 5, tck = 0, 
       col = "black", lty = 1, lwd = 2, mgp = c(0, 0.5, 0),
       cex.axis=2, font = 2)
  axis(2, pos = 0, tcl = 0.2,
       at = seq(round(min(LA$Sr),3), 
                round(max(LA$Sr),3), 0.001), 
       labels = seq(round(min(LA$Sr)*1000,0), 
                    round(max(LA$Sr)*1000,0), 1),
       las = 2, mgp = c(0, 0.5, 0), cex.axis = 2, font = 2, 
       lwd = 2, col = "black", col.axis = "black")
  title(ylab = list("Sr/Ca ratio (mg/g)", cex = 2, font=2),
        xlab = list("distance from core [um]", cex = 2, font=2), 
        col.lab = "black", line = 1)
  text(x = 300, y = 0.008, labels = paste(IDOto[i], "-", 
                                         LA[which(LA$ID == IDOto[i]), "capt",]), 
       cex=3, font=2)

  plot(d18O[which(d18O$ID == IDOto[i]), "sims_dist",],
       d18O[which(d18O$ID == IDOto[i]), "d18O",], main = "",
       xlab =  "", ylab = "",
       xlim = c(0, max(d18O[which(d18O$ID == IDOto[i]), "sims_dist",], na.rm = T)),
       ylim = c(min(d18O$d18O, na.rm = T), 0),
       col.axis = "black", cex.axis = 1, type = "b", col = "blue", lwd = 3,
       fg = "black", cex = 1, axes = FALSE, pch = 19)
  axis(1, pos = mean(d18O[which(d18O$ID == IDOto[i]), "d18O",], na.rm = T),
       at = c(0, 200, max(d18O[which(d18O$ID == IDOto[i]), "sims_dist",], na.rm = T)),
       labels = c("", "indiv.mean", ""), lwd.ticks = 0, tck = 0,
       col = "black", lty = 3, lwd = 1, mgp = c(0,0,0))
  axis(1, pos = mean(d18O$d18O, na.rm = T),
       at = c(0, 200, max(d18O$sims_dist)),
       labels = c("", "global mean", ""), lwd.ticks = 0, tck = 0,
       col = "black", lty = 1, lwd = 1, mgp = c(0,0,0))
  axis(1, pos = round(min(d18O$d18O), 1), at = seq(0, max(d18O$sims_dist), 500),
       labels = seq(0, max(d18O$sims_dist), 500), lwd.ticks = 5, tck = 0,
       col = "black", lty = 1, lwd = 2, mgp = c(0, 0.5, 0))
  axis(2, pos = 0, tcl = 0.2,
       at = seq(round(min(d18O$d18O, na.rm = T),1),
                round(max(d18O$d18O, na.rm = T),1), 1),
       labels = seq(round(min(d18O$d18O, na.rm = T),1),
                    round(max(d18O$d18O, na.rm = T),1), 1),
       las = 2, mgp = c(0, 0.5, 0), cex.axis = 1,
       lwd = 2, col = "black", col.axis = "black")
  title(ylab = "d18O", cex.lab = 1,
        col.lab = "black", line = 5)
  title(xlab = "Distance [um]", cex.lab = 1,
        col.lab = "black", line = 1)
  title(main = "d18O")
  text(x = 300, y = 0, labels = d18O[which(d18O$ID == IDOto[i]), "capt",])
  dev.off()  
}
```

```{r lifelong niche, include=FALSE, eval=FALSE}
#script to form annual means of elemental data for clustering and increment analysis
rm(list = ls())
#Read in data
oto.int <- fread("Data/pike/elements/Oto_LA-to-d18O_interpolation_Sr-corrected_smooth.csv")

oto.int.lifehist <- oto.int%>%
  filter(area!="Control")%>%
  transmute(ID=ID,
            age=as.factor(age),
            capt=capt,
            area=case_when(area=="WRB"~"Lagoon",
                           area=="NRB"~"Lagoon",
                           area=="GB"~"Lagoon",
                           area=="Freshwater"~"Tributary"),
            d18O=d18O,
            d18O_res=res,
            Sr=Sr,
            year=as.factor(year),
            dist=dist)

# Plot lifelong increase in d18O --> colder the older
mypallette <- viridis::viridis(4, direction = 1)

d18O <- ggplot(oto.int.lifehist, aes(year, d18O_res))+
  geom_boxplot(aes(color = area), width = 0.5, outlier.size = .5)+
  scale_x_discrete(limits = factor(1:10),
                   labels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"))+
  labs(x="Age (years)",
       y=expression("Sr-corr."~delta^18*"O"~("VPDB \u2030")),
       color="area of capture")+
  theme_classic()+
  scale_color_viridis_d(direction = 1, begin = 0, end = 0.7)+
  theme(panel.grid.major = element_line(),
        legend.position = "none",
        legend.background = element_blank(),
        legend.box.background = element_blank(),
        axis.line = element_line(size = .5, colour = "black", linetype = 1),
        axis.ticks = element_line(size = .5, colour = "black"),
        axis.text = element_text(size = 8, colour = "black"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8,
                                    margin = margin(r=15)),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key.size = unit(.8,"line"))+
  coord_cartesian(clip = "off")+
  geom_segment(x=-0.56, y=0, xend=-0.56, yend=2.5, 
               lineend = "butt", linejoin = "mitre",
               size=3, arrow = arrow(length = unit(0.2, "inches")),
               color="#66CCCC")+
  annotation_custom(grob = text_grob(label = "lower temperature", 
                                     rot = 90, face = "bold", size = 5),
                                     ymin=0, ymax=2.5,
                                     xmin=-0.57, xmax=-0.57)+
  geom_segment(x=-0.56, y=0, xend=-0.56, yend=-2.5, 
               lineend = "butt", linejoin = "mitre",
               size=3, arrow = arrow(length = unit(0.2, "inches")),
               color="#FF6666")+
  annotation_custom(grob = text_grob(label = "higher temperature", 
                                     rot = -90, face = "bold", size = 5),
                                     ymin=0, ymax=-2.5,
                                     xmin=-0.54, xmax=-0.54)

#Lifelong increase in Sr/Ca --> saltier the older
SrCa <- ggplot(oto.int.lifehist, aes(year, Sr*1000))+
  geom_boxplot(aes(color = area), width = 0.5, outlier.size = .5)+
  scale_x_discrete(limits = factor(1:10))+
  labs(x="Age (years)",
       y="Sr/Ca (mg/g)",
       color="area of capture")+
  theme_classic()+
  scale_color_viridis_d(direction = 1, begin = 0, end = 0.7)+
  theme(panel.grid.major = element_line(),
        legend.position = "none",
        legend.background = element_blank(),
        legend.box.background = element_blank(),
        axis.line = element_line(size = .5, colour = "black", linetype = 1),
        axis.ticks = element_line(size = .5, colour = "black"),
        axis.text = element_text(size = 8, colour = "black"),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.key.size = unit(.8,"line"))

allplots <- ggarrange(d18O, SrCa, ncol = 2, nrow = 1, labels = c("A", "B"), common.legend = TRUE, legend = "bottom")

ggsave("Figures/supplement/Lifelong_niche.png", allplots, width = 15, height = 6, units = "cm", dpi = 600, bg = "white")
```

## DTW-clustering of otolith timeseries
```{r subset data, include=FALSE, eval=FALSE}
rm(list = ls())
#read in data
oto.int <- fread("Data/pike/elements/Oto_LA-to-d18O_interpolation_Sr-corrected_smooth.csv")

#remove NAs
oto.int <- na.omit(oto.int)

#subset according to cluster
Fgroup <- subset(oto.int, area=="Freshwater")
Bgroup <- oto.int%>%filter(area=="WRB"|area=="NRB"|area=="GB")

#Test length, needs to equal N=120
length(unique(Bgroup$fish_ID))+length(unique(Fgroup$fish_ID))

#data package
fwrite(Fgroup, "Data/pike/clustering/Tributary-sample-transects.csv")
fwrite(Bgroup, "Data/pike/clustering/Lagoon-sample-transects.csv")
```

```{r dtw data prep, include=FALSE, eval=FALSE}
rm(list = ls())

#read in data
Fgroup <- as.data.frame(fread("Data/pike/clustering/Tributary-sample-transects.csv"))
Bgroup <- as.data.frame(fread("Data/pike/clustering/Lagoon-sample-transects.csv"))

#load funtions
source("funcs.R")

#Look for sample size
length(unique(Fgroup$fish_ID))
length(unique(Bgroup$fish_ID))

#Mean transect length
median(table(Fgroup$fish_ID)) #Median length 35.5 data points
median(table(Bgroup$fish_ID)) #Median length 31.5 data points

#Make lifeyear a factor
Bgroup$lifeyear <- as.factor(Bgroup$year)
Fgroup$lifeyear <- as.factor(Fgroup$year)

#reinterpolate to standard length (based on medians)
dataB1 <- interpol(series = Bgroup, length = 35, element = "Sr")
dataB2 <- interpol(series = Bgroup, length = 35, element = "res")

dataF1 <- interpol(series = Fgroup, length = 35, element = "Sr")
dataF2 <- interpol(series = Fgroup, length = 35, element = "res")

dataB <- cbind(dataB1, dataB2[,2])
dataF <- cbind(dataF1, dataF2[,2])

dataB[,2] <- scale(dataB[,2])
dataF[,2] <- scale(dataF[,2])

#Data Lists for hierarchical clustering
IDOto <- unique(dataB$ID)
Bseries <- list()

for (i in 1:length(IDOto)) {
  m <- as.matrix(dataB[which(dataB$ID==IDOto[i]),c(2,3)])
  Bseries[[length(Bseries)+1]] <- m
}

names(Bseries) <- c(IDOto)

IDOto <- unique(dataF$ID)
Fseries <- list()

for (i in 1:length(IDOto)) {
  m <- as.matrix(dataF[which(dataF$ID==IDOto[i]),c(2,3)])
  Fseries[[length(Fseries)+1]] <- m
}

names(Fseries) <- c(IDOto)

#save data series
save(Fgroup,Bgroup,dataB, dataF, Bseries, Fseries, interpol,
     file = "Data/Pike/Pike_element_transects_hierarchical-prep.RData")
```

```{r Hierarchical dtw clustering with even lengths, include=FALSE, eval=FALSE}
rm(list = ls())

#load hierarchical data prep
load("Data/Pike/Pike_element_transects_hierarchical-prep.RData")

#cluster Bodden with cutoff test
clust_B <- tsclust(series = Bseries, type = "hierarchical",
                  k = 2L:6L,centroid = dba,
                  control = hierarchical_control(method = "ward.D"),
                  seed = 5L,args = tsclust_args(dist = list(window.size =5L)))

names(clust_B) <- paste0("k_",2L:6L)

EvalB <- data.frame(sapply(clust_B, cvi, type = "internal"))
EvalB$win <- c("max", "max","max","min","min","max","min")

#cluster FW with cutoff test
clust_F <- tsclust(series = Fseries, type = "hierarchical",
                  k = 2L:6L,centroid = dba,
                  control = hierarchical_control(method = "ward.D"),
                  seed = 101L,args = tsclust_args(dist = list(window.size =5L)))

names(clust_F) <- paste0("k_",2L:6L)

EvalF <- data.frame(sapply(clust_B, cvi, type = "internal"))
EvalF$win <- c("max", "max","max","min","min","max","min")

#Plot best solution lagoons
Lagoonplot <- plot(clust_B$k_4, type="sc", lwd=2)+
  ylim(-3,3)+ggtitle(expression("Sr/"~delta^18~O~"cluster centroids"))+
  labs(x="transect cell", y="scaled value")+
  theme(axis.text = element_text(size = 25, color="black"),
        axis.title = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5))

#Plot best solution Freshwater
FWplot <- plot(clust_F$k_6, type="sc", linetype="solid", cex=2)+
  ylim(-3,3)+ggtitle(expression("Sr/"~delta^18~O~"cluster centroids"))+
  labs(x="transect cell", y="scaled value")+
  theme(axis.text = element_text(size = 25, color="black"),
        axis.title = element_text(size = 25),
        plot.title = element_text(size = 25, hjust = 0.5))

save(Fgroup,Bgroup,dataB, dataF, Bseries, Fseries, interpol, clust_B, clust_F, Lagoonplot, FWplot,
     file = "Data/Pike/Adults_hierarchical_clusters.RData")
```

```{r cluster assignment hierarchical, include=FALSE, eval=FALSE}
rm(list=ls())
load("Data/pike/clustering/Adults_element_transects_hierarchical_all.RData")
#assign clusters to fish
Lagoonplot

fishassign_B2 <- data.frame(ID=names(Bseries), 
                            clust2=factor(clust_B$k_4@cluster, 
                                           levels = c(1:4), 
                                           labels = c("BW int saline|cold",
                                                      "BW Low saline|shift",
                                                      "BW high saline|warm",
                                                      "BW high saline|cold")))

FWplot

fishassign_F2 <- data.frame(ID=names(Fseries), 
                            clust2=factor(clust_F$k_6@cluster, 
                                           levels = c(1:6), 
                                           labels = c("FW low saline|shift",
                                                      "FW low saline|cold",
                                                      "FW int saline|const",
                                                      "FW high saline|shift",
                                                      "FW int saline|shift",
                                                      "FW high saline|const")))
#Join dtw clusters with pike data
fishassign_B <- Bgroup%>%
  group_by(ID)%>%
  transmute(ID=ID,
            capt=capt,
            area=area)%>%
  distinct()%>%
  inner_join(fishassign_B2, by="ID")%>%
  mutate(clust=fct_relevel(clust2,
                           "BW Low saline|shift",
                           "BW int saline|cold",
                           "BW high saline|warm",
                           "BW high saline|cold"))

fishassign_F <- Fgroup%>%
  group_by(ID)%>%
  transmute(ID=ID,
            capt=capt,
            area=area)%>%
  distinct()%>%
  inner_join(fishassign_F2, by="ID")%>%
  mutate(clust2=fct_relevel(clust2,
                            "FW low saline|cold",
                            "FW low saline|shift",
                            "FW int saline|const",
                            "FW int saline|shift",
                            "FW high saline|const",
                            "FW high saline|shift"))

fwrite(fishassign_B, file = "Data/Pike/clustering/Bcluster_all.csv")
fwrite(fishassign_F, file = "Data/Pike/clustering/Fcluster_all.csv")


save(Fgroup,Bgroup,dataB, dataF, Bseries, Fseries, interpol, clust_B, clust_F,
     fishassign_B, fishassign_F, Lagoonplot, FWplot,
     file = "Data/Pike/Pike_hierarchical_clusters.RData")
```

```{r table of poportions, include=FALSE, eval=FALSE}
rm(list = ls())

Groupall <- fread("Data/Pike/clustering/Ecotypes-element-transects-all.csv")

allpike_eco <- Groupall%>%
  mutate(ecotype=case_when(clust2=="FW low saline|cold"~"FW-resident",
                              clust2=="FW low saline|shift"~"FW-resident",
                              clust2=="FW int saline|const"~"FW-resident",
                              clust2=="FW int saline|shift"~"Anadromous",
                              clust2=="FW high saline|shift"~"Anadromous",
                              clust2=="FW high saline|const"~"Trans-habitat",
                              clust2=="BW Low saline|shift"~"Trans-habitat",
                              clust2=="BW int saline|cold"~"BW-resident",
                              clust2=="BW high saline|cold"~"BW-resident",
                              clust2=="BW high saline|warm"~"BW-resident"))%>%
  transmute(ID=ID, 
            area=area.x,
            phenotype=ecotype,
            clust=clust2)%>%
  distinct()

Clust_props_per_area <- table(allpike_eco$clust)

Pheno_props_per_area <- table(allpike_eco$area, allpike_eco$phenotype)
write.table(Pheno_props_per_area, "Data/Pike/clustering/Lagoonprobs.csv", sep = ",")
```

```{r plot freshwater clusters, include=FALSE, eval=FALSE}
rm(list=ls())
load("Data/pike/clustering/Adults_element_transects_hierarchical_all.RData")

Bgroup.clus <- Bgroup%>%inner_join(fishassign_B, by="ID")
Fgroup.clus <- Fgroup%>%inner_join(fishassign_F, by="ID")

Groupall <- rbind(Bgroup.clus, Fgroup.clus)

agebreaks <- Groupall%>%group_by(ID, year)%>%
  summarise(yeardist=max(dist))%>%
  ungroup()%>%group_by(year)%>%
  mutate(mean.yeardist=mean(yeardist))

agebreaks.num <- unique(agebreaks$mean.yeardist)
agebreaks.label <- as.character(seq(1,11,1))


# Generate a color ramp between two colors
mypallette <- c(colorRampPalette(c("#004643", "#F1C40F"))(4))

A <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW low saline|cold",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[4], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,-15,-10,10)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "black", size = 7),
        axis.ticks.x = element_line(colour = "transparent"),
        axis.ticks.y = element_line(size = 0.3),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3))+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "FW low saline|cold (N = 4)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = -50, xmax = -50)+
  geom_segment(x=-450, y=-0.5, xend=-450, yend=8.5, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="darkorange")+
  annotation_custom(grob = text_grob(label = "increasing salinity", 
                                     rot = 90, size = 7),
                                     ymin=0, ymax=8,
                                     xmin=-460, xmax=-460)

A1 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW low saline|cold",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[4], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,-15,-5,5.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "black", size = 7),
        axis.ticks.y = element_line(size = 0.3),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3))+
  coord_cartesian(clip = "off")+
    geom_segment(x=-450, y=0, xend=-450, yend=-3.2, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="#FF6666")+
  annotation_custom(grob = text_grob(label = "warmer", 
                                     rot = 90,, size = 7),
                                     ymin=0, ymax=-3,
                                     xmin=-460, xmax=-460)+
  geom_segment(x=-450, y=0, xend=-450, yend=3, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="#66CCCC")+
  annotation_custom(grob = text_grob(label = "colder", 
                                     rot = 90, size = 7),
                                     ymin=0, ymax=3.2,
                                     xmin=-460, xmax=-460)

B <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW low saline|shift",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[1], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,-5,-10,-6)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "FW low saline|shift (N = 9)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

B1 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW low saline|shift",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[1], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,-5,-5,-10)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

C <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW int saline|const",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[2], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,5,-10,-16)), 
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "FW int saline|const (N = 14)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

C1 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW int saline|const",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[2], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,5,-5,-20.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

D <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW int saline|shift",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression(""))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,15,-10,-20)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.line = element_line(size = 0.3),
        axis.title.x = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "FW int saline|shift (N = 7)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

D1 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW int saline|shift",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,15,-5,-31)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

E <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW high saline|shift",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression(""))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,25,-10,-30)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.line = element_line(size = 0.3),
        axis.title.x = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "FW high saline|shift (N = 9)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

E1 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW high saline|shift",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,25,-5,-41.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

F1 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW high saline|const",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression(""))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,35,-10,-40)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.line = element_line(size = 0.3),
        axis.title.x = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "FW high saline|const (N = 11)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

F2 <- ggplot(Fgroup.clus[Fgroup.clus$clust2=="FW high saline|const",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,35,-5,-51.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

Transects <- ggarrange(A, B, C, D, E, F1, A1, B1, C1, D1, E1, F2, labels = c("A", "B", "C", "D", "E", "F"), 
                       hjust = c(-3.85,-2.6,-1.2,0.2,1.6,3.4), vjust = c(2,2,2,2,2), 
                       font.label = list(size=10), nrow = 2, ncol = 6)

Transects2 <- annotate_figure(Transects, bottom = text_grob(expression("Distance from Core"~(mu~"m")), vjust = 1, size = 7))

ggsave("Figures/supplement/FW-clusters-life-history-transects.png", Transects2, dpi = 600, width = 24, height = 8, units = "cm", bg = "white")
```

```{r plot lagoon clusters, include=FALSE, eval=FALSE}
rm(list=ls())
load("Data/pike/clustering/Adults_element_transects_hierarchical_all.RData")

Bgroup.clus <- Bgroup%>%inner_join(fishassign_B, by="ID")
Fgroup.clus <- Fgroup%>%inner_join(fishassign_F, by="ID")

Groupall <- rbind(Bgroup.clus, Fgroup.clus)

agebreaks <- Groupall%>%group_by(ID, year)%>%
  summarise(yeardist=max(dist))%>%
  ungroup()%>%group_by(year)%>%
  mutate(mean.yeardist=mean(yeardist))

agebreaks.num <- unique(agebreaks$mean.yeardist)
agebreaks.label <- as.character(seq(1,11,1))


# Generate a color ramp between two colors
mypallette <- c(colorRampPalette(c("#004643", "#F1C40F"))(4))

A <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW Low saline|shift",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[4], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,-15,-10,10)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "black", size = 7),
        axis.ticks.x = element_line(colour = "transparent"),
        axis.ticks.y = element_line(size = 0.3),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3))+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "BW Low saline|shift (N = 20)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = -50, xmax = -50)+
  geom_segment(x=-450, y=-0.5, xend=-450, yend=8.5, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="darkorange")+
  annotation_custom(grob = text_grob(label = "increasing salinity", 
                                     rot = 90, size = 7),
                                     ymin=0, ymax=8,
                                     xmin=-460, xmax=-460)

A1 <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW Low saline|shift",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[4], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,-15,-5,5.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "black", size = 7),
        axis.ticks.y = element_line(size = 0.3),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3))+
  coord_cartesian(clip = "off")+
    geom_segment(x=-450, y=0, xend=-450, yend=-3.2, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="#FF6666")+
  annotation_custom(grob = text_grob(label = "warmer", 
                                     rot = 90,, size = 7),
                                     ymin=0, ymax=-3,
                                     xmin=-460, xmax=-460)+
  geom_segment(x=-450, y=0, xend=-450, yend=3, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="#66CCCC")+
  annotation_custom(grob = text_grob(label = "colder", 
                                     rot = 90, size = 7),
                                     ymin=0, ymax=3.2,
                                     xmin=-460, xmax=-460)

B <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW int saline|cold",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[1], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,-5,-10,-6)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "BW int saline|cold (N = 17)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

B1 <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW int saline|cold",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[1], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,-5,-5,-10)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

C <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW high saline|warm",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[2], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,5,-10,-16)), 
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "BW high saline|warm (N = 22)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

C1 <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW high saline|warm",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[2], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,5,-5,-20.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

D <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW high saline|cold",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression(""))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,15,-10,-20)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.line = element_line(size = 0.3),
        axis.title.x = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "BW high saline|cold (N = 7)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

D1 <- ggplot(Bgroup.clus[Bgroup.clus$clust2=="BW high saline|cold",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypallette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,15,-5,-31)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

Transects <- ggarrange(A, B, C, D, A1, B1, C1, D1, labels = c("A", "B", "C", "D"), hjust = c(-3.85,-2.6,-1.2,0.15), vjust = c(2,2,2,2), font.label = list(size=10), nrow = 2, ncol = 4)

Transects2 <- annotate_figure(Transects, bottom = text_grob(expression("Distance from Core"~(mu~"m")), vjust = 1, size = 7))

ggsave("Figures/supplement/BW-clusters-life-history-transects.png", Transects2, dpi = 600, width = 16, height = 8, units = "cm", bg = "white")
```

```{r plot life history phenotypes, include=FALSE, eval=FALSE}
rm(list=ls())
Groupall <- fread("Data/Pike/clustering/Ecotypes-element-transects-all.csv")

mypalette <- adjustcolor(c("#004643","#00BFB2","#F1C40F","#C5D86D","#AF3800"), alpha.f = 0.6)

allpike_eco <- Groupall%>%
  mutate(ecotype=case_when(clust2=="FW low saline|cold"~"FW-resident",
                              clust2=="FW low saline|shift"~"FW-resident",
                              clust2=="FW int saline|const"~"Anadromous",
                              clust2=="FW int saline|shift"~"Anadromous",
                              clust2=="FW high saline|shift"~"Anadromous",
                              clust2=="FW high saline|const"~"Trans-habitat",
                              clust2=="BW Low saline|shift"~"Trans-habitat",
                              clust2=="BW int saline|cold"~"BW-resident",
                              clust2=="BW high saline|cold"~"BW-resident",
                              clust2=="BW high saline|warm"~"BW-resident"))

A <- ggplot(allpike_eco[allpike_eco$ecotype=="FW-resident",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypalette[4], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,-15,-10,10)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "black", size = 7),
        axis.ticks.x = element_line(colour = "transparent"),
        axis.ticks.y = element_line(size = 0.3),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3))+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "freshwater resident (N = 27)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = -50, xmax = -50)+
  geom_segment(x=-450, y=-0.5, xend=-450, yend=8.5, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="darkorange")+
  annotation_custom(grob = text_grob(label = "increasing salinity", 
                                     rot = 90, size = 7),
                                     ymin=0, ymax=8,
                                     xmin=-460, xmax=-460)

A1 <- ggplot(allpike_eco[allpike_eco$ecotype=="FW-resident",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypalette[4], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,-15,-5,5.5)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "black", size = 7),
        axis.ticks.y = element_line(size = 0.3),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3))+
  coord_cartesian(clip = "off")+
    geom_segment(x=-450, y=0, xend=-450, yend=-3.2, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="#FF6666")+
  annotation_custom(grob = text_grob(label = "warmer", 
                                     rot = 90,, size = 7),
                                     ymin=0, ymax=-3,
                                     xmin=-460, xmax=-460)+
  geom_segment(x=-450, y=0, xend=-450, yend=3, 
               lineend = "butt", linejoin = "mitre",
               size=2, arrow = arrow(length = unit(0.18, "inches")),
               color="#66CCCC")+
  annotation_custom(grob = text_grob(label = "colder", 
                                     rot = 90, size = 7),
                                     ymin=0, ymax=3.2,
                                     xmin=-460, xmax=-460)

B <- ggplot(allpike_eco[allpike_eco$ecotype=="BW-resident",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypalette[1], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,15,-10,-25.5)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "brackish resident (N = 46)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

B1 <- ggplot(allpike_eco[allpike_eco$ecotype=="BW-resident",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypalette[1], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,15,-5,-30)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

C <- ggplot(allpike_eco[allpike_eco$ecotype=="Anadromous",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypalette[2], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression("Sr/Ca (mg/g)"))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,-5,-10,-6)), 
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "anadromous type (N = 16)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

C1 <- ggplot(allpike_eco[allpike_eco$ecotype=="Anadromous",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypalette[2], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,-5,-5,-10)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")

D <- ggplot(allpike_eco[allpike_eco$ecotype=="Trans-habitat",])+
  geom_line(aes(dist, Sr*1000, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, Sr*1000), color="black", fill=mypalette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(0, 8)+
  scale_y_continuous(breaks = seq(0,8,1), limits = c(0,8))+
  xlim(0,2100)+
  labs(y=expression(""))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(20,5,-10,-9.5)),
        axis.text.x = element_text(color = "transparent", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3, colour = "transparent"),
        axis.title.y = element_text(size = 7, margin = margin(r=10)),
        axis.line = element_line(size = 0.3),
        axis.title.x = element_blank(),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")+
  annotation_custom(grob = text_grob(label = "cross-habitat type (N = 31)", size = 7, just = "left"),
                    ymin = 9.3, ymax = 9.3, xmin = 100, xmax = 100)

D1 <- ggplot(allpike_eco[allpike_eco$ecotype=="Trans-habitat",])+
  geom_line(aes(dist, d18O_res, group=ID), color="grey", linewidth=.3, alpha=.8)+
  geom_smooth(aes(dist, d18O_res), color="black", fill=mypalette[3], 
              alpha=1, size=.3, method = "gam")+
  theme_classic()+
  ylim(-3, 3)+
  xlim(0,2100)+
  labs(y=expression("scaled"~delta^18*"O"~("VPDB \u2030")))+
  theme(legend.position = "none", 
        panel.grid.major.y = element_line(size = .3),
        panel.border = element_rect(fill = NA, size = .3),
        plot.margin = margin(c(10,5,-5,-20)),
        axis.text.x = element_text(color = "black", size = 7),
        axis.text.y = element_text(color = "transparent", size = 7),
        axis.ticks.y = element_line(size = 0.3, colour = "transparent"),
        axis.ticks.length.y = unit(0.05, "cm"),
        axis.ticks.x = element_line(size = 0.3),
        axis.title.y = element_text(size = 7, margin = margin(r=10), colour = "transparent"),
        axis.title.x = element_blank(),
        axis.line = element_line(size = 0.3),
        plot.background = element_blank())+
  coord_cartesian(clip = "off")


Transects <- ggarrange(A, C, D, B, A1, C1, D1, B1, labels = c("A", "B", "C", "D"), hjust = c(-3.85,-2.6,-1.2,0.15), vjust = c(2,2,2,2), font.label = list(size=10), nrow = 2, ncol = 4)

Transects2 <- annotate_figure(Transects, bottom = text_grob(expression("Distance from Core"~(mu~"m")), vjust = 1, size = 7))

ggsave("Figures/MS/Ecotypes-life-history-transects.png", Transects2, dpi = 600, width = 16, height = 8, units = "cm", bg = "white")
```

```{r dissimilarity test, include=FALSE, eval=FALSE}
rm(list = ls())
#read in data
load("Data/Pike/Pike_element_transects_hierarchical_all.RData")

dataLagoon <- Bgroup%>%
  transmute(ID=ID,
            Sr=Sr,
            d18O_res=d18O_res,
            year=year)

dataFW <- Fgroup%>%
  transmute(ID=ID,
            Sr=Sr,
            d18O_res=d18O_res,
            year=year)

#transform data to all positive
dataLagoon$d18O_res <- (dataLagoon$d18O_res+sqrt((min(dataLagoon$d18O_res))^2))
#So a value of d18O res is now to be interpreted as residual from predicted
#value + 2.518462

dataLagoon$Sr <- (dataLagoon$Sr+sqrt((min(dataLagoon$Sr))^2))
#So a value of Sr is now to be interpreted as scaled value + 2.716513

#Same for reshwater group
dataFW$d18O_res <- (dataFW$d18O_res+sqrt((min(dataFW$d18O_res))^2))
dataFW$Sr <- (dataFW$Sr+sqrt((min(dataFW$Sr))^2))

#join
data <- rbind(dataLagoon, dataFW)

#make list of matrices Bodden
IDOto <- unique(data$ID)
series <- list()

for (i in 1:length(IDOto)) {
  m <- as.matrix(data[which(data$ID==IDOto[i]),])
  series[[length(series)+1]] <- m
}

names(series) <- c(IDOto)

#Revert from list to data frame, assign cell numbers
Series <- do.call(rbind.data.frame,series)
Series$ID <- substr(rownames(Series), 1,8)
rownames(Series) <- NULL

#Make cluster list
Bclust <- fishassign_B%>%
  transmute(ID=ID,
            clust=clust2)
Fclust <- fishassign_F%>%
  transmute(ID=ID,
            clust=clust2)

clust <- rbind(Bclust, Fclust)

#Assign years based on cell numbers, join with cluster ID
Series1 <- Series%>%
  inner_join(clust, by="ID")%>%
  transmute(ID=ID,
            year=case_when(year<2~"orig",
                           year>1&year<3~"juv",
                           year>2~"adult"),
            clust=clust,
            Sr=Sr,
            d18O_res=d18O_res)%>%
  group_by(ID, year)%>%
  summarize(ID=ID,
            year=year,
            clust=clust,
            Sr=mean(as.numeric(Sr)),
            d18O_res=mean(as.numeric(d18O_res)))%>%
  distinct()

#Make wide for testing
Series2 <- pivot_wider(Series1, names_from = year, values_from = c(Sr, d18O_res))

#cluster needs to be a factor
Series2$clust <- as.factor(Series2$clust)

#make mvabund object of Sr and d18O data
elements <- mvabund(Series2[,3:8])

boxplot(elements, na.rm=T)

Pikedist <- vegdist(elements, method = "euclidean", na.rm = T)
Disp <- betadisper(Pikedist, Series2$clust)
anova(Disp)

plot(Disp)

adonis2(elements~Series2$clust, method = "euclidean", permutations = 9999, na.rm=T)
```

```{r jackknife reclassification, eval=FALSE}
set.seed(128)

Series3 <- na.omit(Series2)

Series3 <- Series3%>%
  mutate(clust=case_when(clust=="FW low saline|cold"~"FW_resident",
                            clust=="FW low saline|shift"~"FW_resident",
                            clust=="FW int saline|const"~"Anadromous",
                            clust=="FW int saline|shift"~"Anadromous",
                            clust=="FW high saline|shift"~"Anadromous",
                            clust=="FW high saline|const"~"Trans-habitat",
                            clust=="BW Low saline|shift"~"Trans-habitat",
                            clust=="BW int saline|cold"~"BW-resident",
                            clust=="BW high saline|cold"~"BW-resident",
                            clust=="BW high saline|warm"~"BW-resident"))

#jackknife reclassification
lda.m2 <- lda(clust~Sr_orig+Sr_juv+Sr_adult+d18O_res_orig+d18O_res_juv+d18O_res_adult, 
              data = Series3, CV=TRUE)

Series3$jacknife <- lda.m2$class
summary <- Series3 %>% summarise(clust=clust,
                      jackknife=jacknife,
                      lda.error=mean(clust != jacknife))
```

# Genotypes
##Matching phenotype to genotype
```{r genetics data, include=FALSE, eval=FALSE}
rm(list = ls())

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

geneticsdata <- read.delim("Data/pike/genetics/struct_K3_1_f_clean.txt", sep = "", dec = ".", header = F, stringsAsFactors = F)

#1=Freshwater; 2=brackish; 3=anadromous

Data <- geneticsdata%>%
  transmute(ID=paste0(sub("_","-",substrRight(V2,8))),
            cluster1=as.numeric(V6),
            cluster2=as.numeric(V7),
            cluster3=as.numeric(V8))
  

Otodata <- read.delim("Data/Pike/Pikedata.txt", header = T, stringsAsFactors = F, sep = "\t", dec = ".")

Otoclusters <- Otodata%>%
  transmute(ID=ID,
            clust=clust,
            ecotype=ecotype,
            waterbody=capt)%>%
  distinct()

Otogen <- Data%>%left_join(Otoclusters, by="ID")%>%
  drop_na()

new_df <- Otogen %>%
  pivot_longer(cols = starts_with("cluster"), 
               names_to = "cluster", 
               values_to = "value")

# Reshape the data frame to long format
df_long <- pivot_longer(Otogen, -c(ID, ecotype, clust, waterbody), names_to = "cluster", values_to = "prob")

# Calculate total probability for each ecotype
df_total <- aggregate(prob ~ ecotype, df_long, sum)

# Merge total probability with original data frame
df_long <- merge(df_long, df_total, by = "ecotype")

# Calculate proportion for each cluster and ecotype
df_long$prop <- df_long$prob.x / df_long$`prob.y`

# Filter data for ecotype "A"
df_long2 <- df_long%>%
  transmute(ecotype=ecotype,
            ID=ID,
            clust=clust,
            waterbody=waterbody,
            cluster=case_when(cluster=="cluster3"~"anadromous",
                              cluster=="cluster2"~"brackish",
                              cluster=="cluster1"~"freshwater"),
            prob.x=prob.x,
            prob.y=prob.y,
            prop=prop)%>%
  mutate(cluster_lvl=fct_relevel(cluster,
                                 "freshwater",
                                 "anadromous",
                                 "brackish"))

df_F <- subset(df_long2, ecotype == "FW_resident")
df_A <- subset(df_long2, ecotype == "Anadromous")
df_T <- subset(df_long2, ecotype == "Trans-habitat")
df_B <- subset(df_long2, ecotype == "BW-resident")

pl_theme1 =	theme_classic() +
	theme(panel.grid.major = element_line(),
		  legend.box.background = element_rect(),
		  axis.line = element_line(size = .5, colour = "black", linetype = 1),
		  axis.ticks = element_line(size = .5, colour = "black"),
		  axis.text.y = element_text(size = 15, colour = "black"),
		  axis.text.x = element_text(size = 15, colour = "transparent", angle = 90),
		  legend.text = element_text(size = 15),
		  axis.title.x = element_blank(),
		  axis.title.y = element_blank(),
		  plot.title = element_text(size = 15, colour = "black"),
		  plot.margin = unit(c(.8,0,-1.4,1), "cm"),
		  legend.position = "none")

pl_theme2 =	theme_classic() +
	theme(panel.grid.major = element_line(),
		  legend.box.background = element_rect(),
		  axis.line = element_line(size = .5, colour = "black", linetype = 1),
		  axis.ticks = element_line(size = .5, colour = "black"),
		  axis.text.y = element_text(size = 15, colour = "transparent"),
		  axis.text.x = element_text(size = 15, colour = "transparent", angle = 90),
		  legend.text = element_text(size = 15),
		  axis.title.x = element_blank(),
		  axis.title.y = element_blank(),
		  plot.title = element_text(size = 15, colour = "black"),
		  plot.margin = unit(c(.8,1,-1.4,0), "cm"),
		  legend.position = "none")

pl_theme3 =	theme_classic() +
	theme(panel.grid.major = element_line(),
		  legend.box.background = element_rect(),
		  axis.line = element_line(size = .5, colour = "black", linetype = 1),
		  axis.ticks = element_line(size = .5, colour = "black"),
		  axis.text.y = element_text(size = 15, colour = "black"),
		  axis.text.x = element_text(size = 15, colour = "black", angle = 90, vjust = .4, hjust = .95),
		  legend.text = element_text(size = 15),
		  axis.title.x = element_blank(),
		  axis.title.y = element_blank(),
		  plot.title = element_text(size = 15, colour = "black"),
		  plot.margin = unit(c(-0.8,0,0,1), "cm"),
		  legend.position = "none")

pl_theme4 =	theme_classic() +
	theme(panel.grid.major = element_line(),
		  legend.box.background = element_rect(),
		  axis.line = element_line(size = .5, colour = "black", linetype = 1),
		  axis.ticks = element_line(size = .5, colour = "black"),
		  axis.text.y = element_text(size = 15, colour = "transparent"),
		  axis.text.x = element_text(size = 15, colour = "black", angle = 90, vjust = .4, hjust = .95),
		  legend.text = element_text(size = 15),
		  axis.title.x = element_blank(),
		  axis.title.y = element_blank(),
		  plot.title = element_text(size = 15, colour = "black"),
		  plot.margin = unit(c(-0.8,1,0,0), "cm"),
		  legend.position = "none")

# Create stacked bar chart with proportions
FW <- ggplot(df_F, aes(x = cluster_lvl, y = prob.x, fill = cluster)) +
  geom_boxplot()+
  stat_summary(fun=mean, geom="point", shape=18, size=5, color="red", fill="red")+
  labs(title = "freshwater resident", x = "genotype", y = "Assignment probability (%)")+
  ylim(0,1)+
  scale_fill_manual(values = c("#00BFB2", "#004643", "#C5D86D"))+
  theme_bw()+
  pl_theme1

Ana <- ggplot(df_A, aes(x = cluster_lvl, y = prob.x, fill = cluster)) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=18, size=5, color="red", fill="red")+
  labs(title = "anadromous", x = "Cluster", y = "Proportion") +
  ylim(0,1)+
  scale_fill_manual(values = c("#00BFB2", "#004643", "#C5D86D"))+
  theme_bw()+
  pl_theme2

Cross <- ggplot(df_T, aes(x = cluster_lvl, y = prob.x, fill = cluster)) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=18, size=5, color="red", fill="red")+
  labs(title = "cross-habitat", x = "genotye", y = "Assignment probability (%)") +
  ylim(0,1)+
  scale_fill_manual(values = c("#00BFB2", "#004643", "#C5D86D"))+
  theme_bw()+
  pl_theme3

BW <- ggplot(df_B, aes(x = cluster_lvl, y = prob.x, fill = cluster)) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=18, size=5, color="red", fill="red")+
  labs(title = "brackish resident", x = "genotype", y = "Proportion") +
  ylim(0,1)+
  scale_fill_manual(values = c("#00BFB2", "#004643", "#C5D86D"))+
  theme_bw()+
  pl_theme4

allplot <- ggarrange(FW, Ana, Cross, BW)
  

allplot2 <- annotate_figure(allplot, bottom = text_grob("putative genotype", color = "black", size = 15),
                left = text_grob("Assignment probability", color = "black", size = 15, rot = 90, vjust = 2))

ggsave("Figures/MS/Ecotypes_genetic_clusters_k3.pdf", allplot2, 
       width = 15, height = 15, units = "cm", dpi = 600, bg = "white")
```

```{r hard cluster assignment}
rm(list = ls())

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

geneticsdata <- read.delim("Data/pike/genetics/struct_K3_1_f_clean.txt", sep = "", dec = ".", header = F, stringsAsFactors = F)
#1=Freshwater; 2=brackish; 3=anadromous

Data <- geneticsdata%>%
  transmute(ID=paste0(sub("_","-",substrRight(V2,8))),
            cluster1=as.numeric(V6),
            cluster2=as.numeric(V7),
            cluster3=as.numeric(V8))
  

Otodata <- read.delim("Data/Pike/Pikedata.txt", header = T, stringsAsFactors = F, sep = "\t", dec = ".")

Otoclusters <- Otodata%>%
  transmute(ID=ID,
            clust=clust,
            ecotype=ecotype,
            waterbody=capt)%>%
  distinct()

Otogen <- Data%>%left_join(Otoclusters, by="ID")%>%
  drop_na()

Otogen$clust <- apply(Otogen[,c(2,3,4)], 1, FUN = which.max)
Otogen$clustval <- apply(Otogen[,c(2,3,4)], 1, FUN = max)

Otogen2 <- Otogen%>%
  mutate(hardclust = case_when(clustval < 0.7 ~ "U",
                            TRUE~as.character(clust)))

Otogen2$hardclust <- fct_relevel(Otogen2$hardclust, "1", "3", "2", "U")

fwrite(Otogen2, file = "Data/pike/genetics/discrete-cluster-assignment.csv")

#Test if phenotype is predicted by genotype
chisq.test(table(Otogen2$ecotype, Otogen2$hardclust))

#Test if genotype is predicted by capture location
chisq.test(table(Otogen2$waterbody, Otogen2$hardclust))

#Figure for hard assignments
Hardclusts <- ggplot(Otogen2, aes(hardclust, fill = ecotype))+geom_bar(position = "fill", width = 0.7)+theme_classic()+
  scale_x_discrete(breaks = c("1","3","2","U"), 
                   labels=c("freshwater", "anadromous", "brackish" ,"unknown"))+
  scale_y_continuous(breaks = seq(0,1,0.2), 
                     labels = c("0,0", "0,2", "0,4", "0,6", "0,8", "1,0"))+
  scale_fill_manual(breaks = c("Trans-habitat", "FW_resident", "Anadromous", "BW-resident"), 
                    labels = c("cross-habitat", "anadromous", "freshwater resident", "brackish resident"),
                    values = c("#F1C40F", "#00BFB2", "#C5D86D", "#004643"))+
  theme(axis.line = element_line(size = .5, colour = "black", linetype = 1),
		  axis.ticks = element_line(size = .5, colour = "black"),
		  axis.text.y = element_text(size = 10, colour = "black"),
		  axis.text.x = element_text(size = 10, colour = "black"),
		  legend.text = element_text(size = 10))+
  labs(x="genotype", y="proportion", fill="ecotype")

ggsave("Figures/supplement/Ecotypes-genetic-clusters-k3_hard-assigned.pdf", Hardclusts, 
       width = 15, height = 9, units = "cm", dpi = 600, bg = "white")
```

```{r MANOVA genetics, include=FALSE, eval=FALSE}
#Test genetic structuring by ecotype
Otogen2 <- Otogen%>%
  transmute(ID=ID,
            ecotype=ecotype,
            cluster1=cluster1,
            cluster2=cluster2,
            cluster3=cluster3)

#Test spatial structuring by genotype
Otogen3 <- Otogen%>%
  transmute(ID=ID,
            waterbody=waterbody,
            cluster1=cluster1,
            cluster2=cluster2,
            cluster3=cluster3)

Otogen2$ecotype <- as.factor(Otogen2$ecotype)
Otogen3$waterbody <- as.factor(Otogen3$waterbody)

Genetics <- mvabund(Otogen2[3:5])

boxplot(Genetics, na.rm=T)

Gendist <- vegdist(Genetics, method = "euclidean", na.rm = T)

#MDS <- metaMDS(Gendist)
PCA <- prcomp(Genetics)

plot(PCA)

fviz_pca_biplot(PCA, repel = TRUE)

plot(PCA$x[,1], PCA$x[,2], col=factor(Otogen$ecotype))

Disp <- betadisper(Gendist, Otogen2$ecotype)
anova(Disp)

plot(Disp)

adonis2(Genetics~Otogen2$ecotype, method = "euclidean", permutations = 9999, na.rm=T)

pairwise.perm.manova(Genetics, Otogen2$ecotype, test = c("Pillai", "Wilks",
  "Hotelling-Lawley", "Roy", "Spherical"), nperm = 999, 
  progress = TRUE, p.method = "fdr", F = FALSE, R2 = FALSE)
```

# Growth analysis of life history phenotypes and genotype
## Age-specific growth
```{r get clean increment growth data, include=FALSE, eval=FALSE}

rm(list = ls())

#read in pike data
Pike1 <- read.delim("C:/Users/timor/Nextcloud2/Boddenhecht/Data/Pike Data/1.pike_data_base_FOR_USERS/1Pike_Latest_Version.txt", 
                    sep = ";", dec = ".", header = T, stringsAsFactors = F)
Pike2 <- Pike1%>%
  transmute(ID=case_when(Fish_ID=="BH-91050.2"~"BH-91050",
                         Fish_ID=="BH-51614"&Waterbody=="BAT"~"NA",
                         Fish_ID=="BH-90307"&MarkorRecap=="Mark"~"NA",
                         TRUE~Fish_ID),
            capt_date=Date,
            TL_mm = Total_Length_mm,
            Weight=Weight,
            sex=Sex,
            area=Waterbody)

#load in age readings
pike_increments_2022 <- read.delim("Data/Pike/Sample_2022_oto_increments.txt", header = T, stringsAsFactors = F)
pike_increments_2021 <- read.delim("Data/Pike/Sample_2021_oto_increments.txt", header = T, stringsAsFactors = F)
pike_increments_pilot <- read.delim("Data/Pike/Pilot_sample_oto_increments.txt", header = T, stringsAsFactors = F)

pike_increments_pilot$Sample <- substr(pike_increments_pilot$Sample,3,10)

pike_increments <- rbind(pike_increments_2021, pike_increments_2022, pike_increments_pilot)

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}

pike_increments2 <- pike_increments%>%
  group_by(Sample)%>%
  transmute(ID=substrRight(Sample,8),
            Quality=Quality,
            Age=Age,
            Cohort=Cohort,
            Core=Core,
            lifeyear=I,
            total_rad_um=sum(Increment, na.rm = T),
            year=Year,
            Increment=Increment)%>%
  ungroup()

pike_inc <- pike_increments2 %>% inner_join(Pike2, by="ID")%>%
  group_by(ID)%>%
  transmute(ID=ID,
            age=Age,
            cohort=Cohort,
            capt=area,
            TL=TL_mm,
            weigth=Weight,
            sex=sex,
            date=lubridate::date(parse_date_time(capt_date, "ymd")),
            core=Core,
            lifeyear=as.integer(lifeyear),
            year=year,
            increment=Increment,
            total_rad_um=total_rad_um,
            quality=Quality)%>%
  drop_na(lifeyear)%>%
  arrange(ID,lifeyear)

pike_inc <- as.data.frame(pike_inc)

#convert inc to rad
IDpike <- unique(pike_inc$ID)
rad <- NULL
for (i in 1:length(IDpike)) {
  d <- cumsum(coalesce(pike_inc[which(pike_inc$ID == IDpike[i]), "increment"], 0)) + pike_inc[which(pike_inc$ID == IDpike[i]), "increment"]*0
  
  rad <- c(rad, d)
}

pike_inc$rad <- rad

pike_inc2 <- pike_inc%>%
  group_by(ID)%>%
  transmute(ID=ID,
            age=age,
            cohort=cohort,
            capt=capt,
            TL_mm=TL,
            weight_g=weigth,
            sex=sex,
            date=date,
            core=core,
            lifeyear=lifeyear,
            year=year,
            increment_um=increment,
            rad_um=rad,
            total_rad_um=total_rad_um,
            quality=quality)%>%
  drop_na(lifeyear)

write.table(pike_inc2, file = "Data/pike/growth/Oto_increment_data_with_rad.txt",
            sep = "\t", dec = ".", row.names = F)
```

```{r join growth measurements with ecotypes, include=FALSE, eval=FALSE}

rm(list = ls())

#read in data on increment widths
pike_inc <- read.delim(file = "Data/Pike/Oto_increment_data_with_rad.txt",
            sep = "\t", dec = ".", header = T, stringsAsFactors = F)

#get clustering results
fishassign_B <- read.delim(file = "Data/Pike/Bcluster_all.txt", 
                           header = T, stringsAsFactors = F, dec = ".", sep = "\t")
fishassign_F <- read.delim(file = "Data/Pike/Fcluster_all.txt", 
                           header = T, stringsAsFactors = F, dec = ".", sep = "\t")

allpike_clust <- rbind(fishassign_B[,c(1,5)], fishassign_F[,c(1,4)])

#join increment data with clustering results
allpike_clust <- allpike_clust%>%
  transmute(ID=ID,
            clust=clust2,
            ecotype=case_when(clust=="FW low saline|cold"~"FW_resident",
                              clust=="FW low saline|shift"~"FW_resident",
                              clust=="FW int saline|const"~"FW_resident",
                              clust=="FW int saline|shift"~"Anadromous",
                              clust=="FW high saline|shift"~"Anadromous",
                              clust=="FW high saline|const"~"Trans-habitat",
                              clust=="BW Low saline|shift"~"Trans-habitat",
                              clust=="BW int saline|cold"~"BW-resident",
                              clust=="BW high saline|cold"~"BW-resident",
                              clust=="BW high saline|warm"~"BW-resident"))

pikedata <- pike_inc%>%
  inner_join(allpike_clust, by = "ID")

write.table(pikedata, "Data/Pike/Growthdata_clusters.txt", row.names = F,
            sep = "\t", dec = ".")
```

```{r include environmental measurements?, include=FALSE, eval=FALSE}
#Script to read cleaned environmental data and combine it with pike data
rm(list = ls())

Lungdata <- read.delim("Data/environment/LUNGdata.txt", header = T, 
                       stringsAsFactors = F, sep = "\t", dec = ".")

pikedata <- read.delim("Data/Pike/Growthdata_clusters.txt", 
                        sep = "\t", dec = ".", header = T, stringsAsFactors = F)


#Clean up data frames
pikedata1 <- pikedata%>%
  transmute(ID=ID,
            area=case_when(capt=="SB"~"WRB",
                           capt=="KB"~"WRB",
                           capt=="Sehrowbach"~"Tributary",
                           capt=="Ziese"~"Tributary",
                           capt=="NHG"~"Tributary",
                           capt=="WB"~"NRB",
                           capt=="BEG"~"NRB",
                           capt=="KJB"~"NRB",
                           capt=="GJB"~"NRB",
                           capt=="GB"~"GB",
                           capt=="Peene"~"Tributary",
                           capt=="Barthe"~"Tributary",
                           capt=="Badendycksgraben"~"Tributary"),
            capt=capt,
            TL_mm=as.numeric(TL_mm),
            weight_g=as.numeric(weight_g),
            sex=as.factor(sex),
            date=date,
            age=age,
            lifeyear=as.numeric(lifeyear),
            year=as.factor(year),
            inc_um=as.numeric(increment_um),
            rad_um=as.numeric(rad_um),
            total_rad_um=as.numeric(total_rad_um),
            clust=as.factor(clust),
            ecotype=as.factor(ecotype))

#Bind together

oto_lowres <- read.delim("Data/Pike/Oto_LA-to-d18O_interpolation_with_res.txt",
                         header = T, dec = ".", sep = "\t",
                         stringsAsFactors = F)

pike_years <- oto_lowres%>%
  group_by(fish_ID,lifeyear)%>%
  summarise(ID=fish_ID,
            lifeyear=as.numeric(lifeyear),
            d18O_mean=mean(d18O, na.rm=T),
            res_mean=mean(res, na.rm=T),
            Sr_mean=mean(Sr, na.rm=T),
            d18O_min=min(d18O, na.rm=T),
            res_min=min(res, na.rm=T),
            Sr_min=min(Sr, na.rm=T),
            d18O_max=max(d18O, na.rm=T),
            res_max=max(res, na.rm=T),
            Sr_max=max(Sr, na.rm=T))%>%
  drop_na()

pikedata2 <- pikedata1%>%left_join(pike_years, by=c("ID", "lifeyear"))%>%
  transmute(ID=ID,
            waterbody=area,
            capt=capt,
            clust=clust,
            ecotype=ecotype,
            age=age,
            weight_g=weight_g,
            TL_mm=TL_mm,
            sex=sex,
            lifeyear=lifeyear,
            year=year,
            inc_um=inc_um,
            rad_um=rad_um,
            res_mean=res_mean,
            res_max=res_max,
            res_min=res_min,
            Sr_mean=Sr_mean,
            Sr_max=Sr_max,
            Sr_min=Sr_min)%>%
  drop_na(inc_um,rad_um,year)%>%
  distinct()

Lungdata$year <- as.factor(Lungdata$year)

pikedata3 <- pikedata2%>%left_join(Lungdata, by=c("waterbody", "year"))

write.table(pikedata3, 
            file = "Data/pike/Pikedata.txt", 
            sep = "\t", dec = ".", row.names = F)
```

```{r join data and cleanup, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
rm(list = ls())

pikes <- read.delim("Data/pike/Pikedata.txt", header = T, stringsAsFactors = F,
                    sep = "\t", dec = ".")

#Inspect data
head(pikes)
tail(pikes)
str(pikes)

#Prepare data frame with factors, scale and center element data
pikes2 <- pikes%>%
  transmute(ID=ID,
            waterbody=as.factor(waterbody),
            TL_mm=TL_mm,
            clust=as.factor(clust),
            phenotype=as.factor(ecotype),
            sex=as.factor(sex),
            lifeyear=lifeyear,
            inc_um=inc_um,
            res_mean=as.numeric(scale(res_mean)),
            Sr_mean=as.numeric(scale(Sr_mean)))%>%
  drop_na(res_mean, Sr_mean)

pikes2[pikes2$ID=="BH-90307",]$sex <- "f"

#Join genotype data
genotypes <- fread("Data/pike/genetics/discrete-cluster-assignment.csv")

genotype_only <- genotypes%>%
  transmute(ID=ID,
            genotype=hardclust)

pikes3 <- pikes2%>%
  inner_join(genotype_only, by = "ID")

fwrite(pikes3, "Data/pike/growth/Pikedata_model.csv")

#Check
str(pikes3)
```

```{r data exploration, include=FALSE, echo=FALSE, eval=FALSE, fig.height=6, fig.width=8}
rm(list = ls())

#read in model data
pikes <- fread("Data/pike/growth/Pikedata_model.csv")

#Exploration script
par(mfrow=c(2,2))
dotchart(pikes2$inc_um, color = as.factor(pikes2$phenotype), main = "Color = Phenotype(Cluster)")
dotchart(pikes2$inc_um, color = as.factor(pikes2$genotype), main = "Color = Genotype(STRUCTURE Cluster)")
dotchart(pikes2$inc, color = as.factor(pikes2$sex), main = "Color = Sex")
dotchart(pikes2$inc, color = as.factor(pikes2$lifeyear), main = "Color = d18O")

par(mfrow=c(3,2))
boxplot(pikes2$inc~pikes2$lifeyear, ylab ="Growth increment", xlab = "Age")
boxplot(pikes2$inc~pikes2$phenotype, ylab ="Growth increment", xlab = "Phenotype")
boxplot(pikes2$inc~pikes2$genotype, ylab ="Growth increment", xlab = "Genotype")
boxplot(pikes2$inc~pikes2$sex, ylab ="Growth increment", xlab = "Sex")
plot(pikes2$inc~pikes2$res_mean, ylab ="Growth increment", xlab = "d18O")
plot(pikes2$inc~pikes2$Sr_mean, ylab ="Growth increment", xlab = "Sr")
plot(0, xaxt = 'n', yaxt = 'n', bty = 'n', pch = '', ylab = '', xlab = '')

par(mfrow = c(2,2))
hist(pikes2$lifeyear^2, main = "", xlab = "Increment (um)")
hist(pikes2$res_mean, main = "", xlab = "d18O ratio")
hist(pikes2$Sr_mean, main = "", xlab = "Sr concentration")
plot(0, xaxt = 'n', yaxt = 'n', bty = 'n', pch = '', ylab = '', xlab = '')
```

Model form
$$log10(Inc) \sim Age*phenotype+Age*genotype+d18Ores+Srmean+turbidity+salinity+WT+P+Sex+(1|Loc/ID)$$

```{r decide age effect, include=FALSE, eval=FALSE}
rm(list = ls())

#read in model data
pikes2 <- fread("Data/pike/growth/Pikedata_model.csv")

#Run model
pikemodel.1 <- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+exp(-lifeyear)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)
pikemodel.1.1 <- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+I(lifeyear^2)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)

anova(pikemodel.1, pikemodel.1.1)
summary(pikemodel.1)
#Quadratic term has lower AIC and residuals look better
```
 
$$log10(Inc) \sim Age*d18Ores+cluster+Age^2+Srmean+Sex+(1|Location)$$

```{r variance inflation factors, echo=TRUE, eval=FALSE}
#Look at variance inflation factors
pikemodel.1<- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+I(lifeyear^2)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)

pike.vif <- vif(pikemodel.1)

barplot(pike.vif[,3], main = "VIF values", horiz = F, col = "darkgreen", ylim = c(0,6),
        las = 2)
abline(h = 5, lwd = 3, lty = 2, col="red")
```

```{r test effects, echo=TRUE, eval=FALSE}
rm(list = ls())

#read in model data
pikes2 <- fread("Data/pike/growth/Pikedata_model.csv")

#run model
pikemodel.1<- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+I(lifeyear^2)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)

#Test random effects
rand(pikemodel.1)
#random effect is significant

#Test age effect
pikemodel.noage <- update(pikemodel.1, ~.-I(lifeyear^2))

anova(pikemodel.1, pikemodel.noage)
#age is significant

#Test life history phenotype effect (interaction term)
pikemodel.nopheno <- update(pikemodel.1, ~.-lifeyear*phenotype)

anova(pikemodel.1, pikemodel.nopheno)
#age*phenotype is significant

#test genotype effect
pikemodel.nogen <- update(pikemodel.1, ~.-lifeyear*genotype)

anova(pikemodel.1, pikemodel.nogen)

#Test thermal effects
pikemodel.nores <- update(pikemodel.1, ~.-res_mean)

anova(pikemodel.1, pikemodel.nores)
#relative d18O (thermal proxy) significant

#Test salinity effects
pikemodel.noSr <- update(pikemodel.1, ~.-Sr_mean)

anova(pikemodel.1, pikemodel.noSr)
#Salinity (Sr:Ca) significant

#Test sex effects
pikemodel.nosex <- update(pikemodel.1, ~.-sex)

anova(pikemodel.1, pikemodel.nosex)
#sex is significant
```

```{r age-specific model summary and table, out.width="120%", dpi=300, echo=FALSE, eval=FALSE, message=FALSE}
rm(list = ls())

#read in model data
pikes2 <- fread("Data/pike/growth/Pikedata_model.csv")

pikemodel.final <- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+I(lifeyear^2)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)

summary(pikemodel.final)
sd(resid(pikemodel.final))
r.squaredGLMM(pikemodel.final)
tab_model(pikemodel.final, show.fstat = T, show.loglik = T, digits.re = 20, show.re.var = T)
```

```{r age-specific model validation, echo=FALSE, eval=FALSE}
rm(list = ls())

#read in model data
pikes2 <- fread("Data/pike/growth/Pikedata_model.csv")

#validation
pikemodel.final <- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+I(lifeyear^2)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)

pikes2$resid <- resid(pikemodel.final)

png(filename = "Figures/supplement/S2_residual_plot.png", width = 5000, height = 2500, res = 600)
plot(pikemodel.final, xlab = "Fitted", ylab = "Residuals")
dev.off()

png(filename = "Figures/supplement/S3_residual_hist.png", width = 5000, height = 2500, res = 600)
hist(pikes2$resid, main = "", xlab = "Residuals")
dev.off()

png(filename = "Figures/supplement/S4_qqnorm.png", width = 5000, height = 2500, res = 600)
qqnorm(resid(pikemodel.final))
qqline(resid(pikemodel.final))
dev.off()

png(filename = "Figures/supplement/S5_params.png", width = 5000, height = 2500, res = 600)
par(mfrow=c(2,3))
plot(pikes2$resid~as.factor(pikes2$phenotype), ylab = "Residuals", xlab = "", las = 2)
plot(pikes2$resid~as.factor(pikes2$genotype), ylab = "Residuals", xlab = "", las = 2)
plot(pikes2$resid~as.factor(pikes2$sex), ylab = "Residuals", xlab = "Sex")
plot(pikes2$resid~pikes2$lifeyear, ylab = "Residuals", xlab = "Age")
plot(pikes2$resid~pikes2$res_mean, ylab = "Residuals", xlab = "d18O")
plot(pikes2$resid~pikes2$Sr_mean, ylab = "Residuals", xlab = "Sr concentration")
dev.off()
```

```{r figures for abiotic main effects, eval=FALSE}
rm(list = ls())

pikes2 <- fread("Data/pike/growth/Pikedata_model.csv")

#factorize temperature
pikes2$agefact <- ifelse(pikes2$lifeyear<3, 2, 
                         ifelse(pikes2$lifeyear>2&pikes2$lifeyear<7, 1,
                                ifelse(pikes2$lifeyear>6, 3, NA)))

pikes2$agefact2 <- factor(pikes2$agefact,levels = c(2,1,3),
                                            labels = c("0-2 years", "3-6 years", 
                                                       "> 6 years"))

mypallette <- viridis::viridis(4, direction = -1)

#Temperature effect
Temp <- ggplot(pikes2, aes(res_mean, log(inc_um), color=agefact2))+
  geom_point(size=.5)+
  stat_smooth(method = "lm", fill="gray", fullrange = T, lwd=.8)+
  labs(x=expression("scaled"~delta^18*"O"~("z-score")), 
       y=expression("log increment"~(mu*"m")),
       color="ageclass")+
  scale_color_viridis_d()+
  theme_classic()+
  theme(panel.grid.major = element_line(),
        legend.position = c(.9,.82),
        legend.background = element_blank(),
        legend.box.background = element_rect(),
        axis.line = element_line(size = .5, colour = "black", linetype = 1),
        axis.ticks = element_line(size = .5, colour = "black"),
        axis.text = element_text(size = 15, colour = "black"),
        axis.title.x = element_text(size = 15, margin = margin(t=15), colour = "black"),
        axis.title.y = element_text(size = 15, colour = "black"),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 15, colour = "black"),
        legend.key.size = unit(.8,"line"))+
  coord_cartesian(clip = "off")+
  geom_segment(x=0, y=3.2, xend=2.5, yend=3.2, 
               lineend = "butt", linejoin = "mitre",
               size=3, arrow = arrow(length = unit(0.18, "inches")),
               color="#66CCCC")+
  annotation_custom(grob = text_grob(label = "colder", 
                                     rot = 0, face = "bold", size = 15),
                                     ymin=3.22, ymax=3.22,
                                     xmin=0, xmax=2.5)+
  geom_segment(x=0, y=3.2, xend=-2.5, yend=3.2, 
               lineend = "butt", linejoin = "mitre",
               size=3, arrow = arrow(length = unit(0.18, "inches")),
               color="#FF6666")+
  annotation_custom(grob = text_grob(label = "warmer", 
                                     rot = -0, face = "bold", size = 15),
                                     ymin=3.22, ymax=3.22,
                                     xmin=0, xmax=-2.5)

ggsave("Figures/MS/Temperature-effect-growth.png", Temp, width = 16, height = 8, units = "cm", dpi = 600, bg = "white")

Sal <- ggplot(pikes2, aes(Sr_mean, log(inc_um), color=agefact2))+
  geom_point(size=.5)+
  stat_smooth(method = "lm", fill="gray", fullrange = T, lwd=.8)+
  labs(x=expression("scaled"~"Sr/Ca"~("z-score")), 
       y=expression("log increment"~(mu*"m")),
       color="ageclass")+
  scale_color_viridis_d()+
  theme_classic()+
  theme(panel.grid.major = element_line(),
        legend.position = c(.9,.92),
        legend.background = element_blank(),
        legend.box.background = element_rect(),
        axis.line = element_line(size = .5, colour = "black", linetype = 1),
        axis.ticks = element_line(size = .5, colour = "black"),
        axis.text = element_text(size = 15, colour = "black"),
        axis.title.x = element_text(size = 15, margin = margin(t=15), colour = "black"),
        axis.title.y = element_text(size = 15, colour = "black"),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 15, colour = "black"),
        legend.key.size = unit(.8,"line"),
        plot.margin = margin(25,1,1,1))+
  coord_cartesian(clip = "off")+
  geom_segment(x=0, y=3.2, xend=2, yend=3.2, 
               lineend = "butt", linejoin = "mitre",
               size=3, arrow = arrow(length = unit(0.18, "inches")),
               color="darkorange")+
  annotation_custom(grob = text_grob(label = "higher salinity", 
                                     rot = 0, face = "bold", size = 15),
                                     ymin=3.22, ymax=3.22,
                                     xmin= 0.5, xmax=1.5)+
  geom_segment(x=0, y=3.2, xend=-2, yend=3.2, 
               lineend = "butt", linejoin = "mitre",
               size=3, arrow = arrow(length = unit(0.18, "inches")),
               color="lightblue")+
  annotation_custom(grob = text_grob(label = "lower salinity", 
                                     rot = -0, face = "bold", size = 15),
                                     ymin=3.22, ymax=3.22,
                                     xmin= -0.5, xmax= -1.5)

ggsave("Figures/MS/Salinity-effect-growth.png", Sal, width = 16, height = 9, units = "cm", dpi = 600, bg = "white")
```

```{r figure for phenotye effect, eval=FALSE}
rm(list = ls())

pikes2 <- fread("Data/pike/growth/Pikedata_model.csv")

pikemodel.final <- lmer(log10(inc_um)~res_mean+lifeyear*phenotype+lifeyear*genotype+I(lifeyear^2)+Sr_mean+sex+(1|ID),
                    REML = F, data = pikes2)

mypalette1 <- adjustcolor(c("#004643","#00BFB2","#F1C40F","#C5D86D","#AF3800"), alpha.f = 0.6)

lifeyear <- unique(pikes2$lifeyear)
phenotype <- unique(pikes2$phenotype)
genotype <- unique(pikes2$genotype)
res_mean <- seq(min(pikes2$res_mean), max(pikes2$res_mean), 0.1)
Sr_mean <- seq(min(pikes2$Sr_mean), max(pikes2$res_mean),0.1)
sex <- unique(pikes2$sex)
ID <- "NA"

Newdata <- expand.grid(lifeyear=lifeyear, phenotype=phenotype, genotype=genotype, res_mean=res_mean, Sr_mean=Sr_mean, sex=sex, ID=ID)
head(Newdata)

Newdata$prediction <- 10^(predict(pikemodel.final, newdata = Newdata, allow.new.levels=TRUE, type="response"))

Newdata2 <- Newdata%>%
  mutate(keep=case_when(phenotype=="FW_resident"&lifeyear<=max(pikes2[pikes2$phenotype=="FW_resident",]$lifeyear)~1,
                   phenotype=="Anadromous"&lifeyear<=max(pikes2[pikes2$phenotype=="Anadromous",]$lifeyear)~1,
                   phenotype=="BW-resident"&lifeyear<=max(pikes2[pikes2$phenotype=="BW-resident",]$lifeyear)~1,
                   phenotype=="Trans-habitat"&lifeyear<=max(pikes2[pikes2$phenotype=="Trans-habitat",]$lifeyear)~1,
                   TRUE~0))%>%
           filter(keep==1)

Newdata2$phenotype <- fct_relevel(Newdata2$phenotype, "FW_resident", "Anadromous", "Trans-habitat", "BW-resident")

# Create ggplot with boxplots and custom position dodge 

Ecogrowth <- ggplot(Newdata2, aes(factor(lifeyear), prediction, fill=phenotype), alpha=0.6)+
  geom_boxplot(lwd=.5, outlier.shape = NA, width=.6, position = position_dodge(width = NULL), color="black")+
  labs(x="Age (years)", y=expression("predicted increment"~(mu*"m")),
       fill="behavioral phenotype")+
  scale_x_discrete(limits=factor(c(1:11)),
                   breaks=c("1","2","3","4","5","6","7","8","9","10","11"))+
  scale_fill_manual(breaks = c("FW_resident","Anadromous","Trans-habitat","BW-resident"),
                    values = c("#C5D86D","#00BFB2","#F1C40F","#004643"),
                    labels = c("freshwater resident", "anadromous","cross-habitat", "brackish resident"))+
  theme_classic()+
  theme(panel.grid.major = element_line(),
        legend.position = c(.715,.79),
        legend.background = element_blank(),
        legend.box.background = element_rect(),
        axis.line = element_line(size = 1, colour = "black", linetype = 1),
        axis.ticks = element_line(size = 1, colour = "black"),
        axis.text = element_text(size = 30, colour = "black"),
        axis.text.y = element_text(size = 30, colour = "black"),
        axis.title.x = element_text(size = 30, colour = "black"),
        axis.title.y = element_text(size = 30, colour = "black"),
        legend.text = element_text(size = 30, colour = "black"),
        legend.title = element_text(size = 30, colour = "black"),
        legend.key.size = unit(2.5,"line"))

ggsave("Figures/MS/Ecotype-effects-growth.png", Ecogrowth, width = 16, height = 8, dpi = 600, bg = "white")
```

## Lifelong growth
```{r data prep for VB modelling, include=FALSE, eval=FALSE}
rm(list = ls())

# load data
pikedata <- read.delim("Data/pike/Pikedata.txt", header = T, stringsAsFactors = F)

# Cleanup
growthdata <- pikedata %>%
  transmute(ID = ID,
            ID2=as.numeric(factor(ID)),
            sex = sex,
            weight_g = weight_g,
            TL_mm = TL_mm,
            rad_mm = rad_um/1000,
            area = waterbody,
            age = age,
            lifeyear = lifeyear,
            ecotype_char=case_when(ecotype=="FW_resident"~"FW-resident",
                                   TRUE~ecotype),
            ecotype_num=case_when(ecotype=="BW-resident"~4,
                              ecotype=="FW_resident"~1,
                              ecotype=="Anadromous"~2,
                              ecotype=="Trans-habitat"~3))

fwrite(growthdata, file = "Data/Pike/growth/Pikedata_Stan.csv")
```

```{r run model and save output, include=FALSE, eval=FALSE}
rm(list = ls())

source("Data/pike/growth/funcs.r")
growthdata <- fread("Data/Pike/growth/Pikedata_Stan.csv")


standata_growth = list(
	Radmm = growthdata$rad_mm, 
	Agei = growthdata$lifeyear,
	N = nrow(growthdata),
	n_fish = as.integer(max(growthdata$ID2)),
	n_eco = as.integer(max(growthdata$ecotype_num)),
	fish_id = as.integer(growthdata$ID2))
fishtable = unique(growthdata[, c("ID2", "ecotype_num")])
fishtable = fishtable[order(fishtable$ID2), ]
standata_growth$eco_id_fish = fishtable$ecotype

stancode_i = ("Data/pike/growth/gm_gamma_inv.stan")
mod_i = stan_model(stancode_i, model_name = "Individual pike growth")

fit_gamma_i = sampling(mod_i, iter = 10000, data = standata_growth, init = init_fun,
					 control = list(adapt_delta = 0.99, max_treedepth = 20), chains = 4, thin = 5)

saveRDS(fit_gamma_i, "Data/pike/growth/fit_gamma_i_final_TR.rds")

p = c("LInf", "tZero", "k", "LInf_eco", "tZero_eco", "k_eco")
samps_a = as.array(fit_gamma_i, pars = p)

png("Figures/supplement/trace.png", w = 1500, h = 1500)
	bayesplot::mcmc_trace(samps_a)
dev.off()
png("Figures/supplement/intervals.png", w = 1500, h = 1500)
	bayesplot::mcmc_intervals(samps_a)
dev.off()


samps = as.matrix(fit_gamma_i, pars = p)
dat = data.frame(
	estimate = apply(samps, 2, median),
	lower = apply(samps, 2, quantile, 0.05),
	upper = apply(samps, 2, quantile, 0.95)
)
tab = cbind(rownames(dat), as.character(signif(dat$estimate, 3)), 
			paste0("(", signif(dat$lower, 3), ", ", signif(dat$upper, 3), ")"))
colnames(tab) = c("Parameter", "Posterior median", "90% CI")
knitr::kable(tab)
```

```{r figure VB curves, include=FALSE, eval=FALSE}
rm(list = ls())

# Ectype codes: 1 = "Brackish resident"; 2 = "Freshwater resident"; 3 = "Anadromous"; 4 = "Cross-habitat"
fit_gamma_i <- readRDS("Data/pike/growth/fit_gamma_i_final_TR.rds")
growthdata <- fread("Data/pike/growth/Pikedata_Stan.csv")

mypallette <- c("#C5D86D","#00BFB2","#F1C40F","#004643")
mylabel <- c("freshwater resident","anadromous","cross-habitat", "brackish resident")

Incpredict = function(pars, Agei) pars[1] * (1 - exp(-pars[2] * (Agei - (-pars[3]))))

pl_theme =	theme_classic() +
  theme(panel.grid.major = element_line(),
        legend.box.background = element_rect(),
        axis.line = element_line(size = .5, colour = "black", linetype = 1),
        axis.ticks = element_line(size = .5, colour = "black"),
        axis.text.x = element_text(size = 18, colour = "black"),
        axis.text.y = element_text(size = 18, colour = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 5, face = "bold"),
        legend.key.size = unit(.5,"line"),
        legend.position = c(0.3,0.8),
        plot.title = element_text(size = 20, colour = "black"))

# x is a stanfit
ecotype_predict = function(x, quants = c(0.5, 0.05, 0.95), names = c("median", "lower", "upper")) {
	pars_eco = c("LInf_eco", "k_eco", "tZero_eco")
	samps = as.matrix(x, pars = pars_eco)

	# list of 4 ecotypes
	# each element is a matrix, 20000 rows (samples) by 15 age increments
	eco_pr = lapply(0:3, \(i) t(apply(samps[, c(1, 5, 9) + i], 1, Incpredict, Agei = 1:15)))

	# convert to a data table summarizing by quantiles
	eco_q = lapply(eco_pr, \(x) {
		res = data.frame(t(apply(x, 2, quantile, quants)))
		colnames(res) = names
		res$Agei = 1:15
		res
	})
	rbindlist(eco_q, idcol = "eco_id")
}

fish_predict = function(x, fish_lookup) {
	pars_fish = c("LInf_i", "k_i", "tZero_i")
	samps_f = as.matrix(x, pars = pars_fish)
	fish_pr = lapply(0:119, \(i) t(apply(samps_f[, c(1, 121, 241) + i], 1, Incpredict, Agei = 1:15)))
	fish_m = lapply(fish_pr, \(x) data.frame(fit = apply(x, 2, median), Agei = 1:15))
	fish_plot = rbindlist(fish_m, idcol = "fish_id")	
	fish_plot$eco_id = fish_lookup$ecotype_num[match(fish_plot$fish_id, fish_lookup$ID2)]
	fish_plot
}

ecotype_curves = function(x, dat) {
	fish_lookup = unique(dat[, c("ID2", "ecotype_num")])
	fish_lookup = fish_lookup[order(fish_lookup$ID2),]
	eco_plot = ecotype_predict(x)
	fish_plot = fish_predict(x, fish_lookup)
	
pl_curves = list()
	for(i in 1:4) {
		pl_curves[[i]] = ggplot(data = eco_plot[eco_id == i]) + 
			geom_line(data = fish_plot[eco_id == i], aes(x = Agei, y = fit, group = fish_id), color = "grey", size = 0.5, alpha = 0.3) + 
			geom_ribbon(aes(x = Agei, ymin = lower, ymax = upper), fill = mypallette[i], alpha = 0.5) + 
			geom_line(aes(x = Agei, y = median), color = mypallette[i], linewidth = 2) + 
		  ggtitle(mylabel[i])+
			geom_point(data = dat[dat$ecotype_num == i, ], 
					   aes(x = lifeyear, y = rad_mm, group = ID), size = 1, alpha = .9, color="black") + 
		  ylim(0,3.8)+scale_x_continuous(breaks = seq(1,15,1), labels = c("1","","3","","5","","7","","9","","11","","13","","15"))+
			pl_theme + labs(x="Age", y="otolith radius (um)")
	}
	pl_curves$nrow = 2
	do.call(grid.arrange, pl_curves)
}

Ecocurves <- ecotype_curves(fit_gamma_i, growthdata)

Ecocurves2 <- ggarrange(Ecocurves)

Ecocurves3 <- annotate_figure(Ecocurves2, 
                              bottom = text_grob("Age (years)", rot = 0, size = 20), 
                              left = text_grob(expression("otolith radius"~(mu~"m")), rot = 90, size = 20))

ggsave("Figures/MS/Individual_growth_revised.png", Ecocurves3, width = 15, height = 15, units = "cm", dpi = 600, bg = "white")
```

```{r lifelong growth analysis evaluation, include=FALSE, eval=FALSE}
rm(list = ls())
#Evaluation of Stan model for individual fish
fit <- readRDS("Data/fit_gamma_i_final_TR.rds")
#write data matrix for plotting
samples <- as.array(fit, pars=c("LInf_eco", "k_eco", "tZero_eco", "phi"))
head(samples)

#write arrays for eval
sample_LInf <- as.array(fit, pars = c("LInf_eco"))
sample_k <- as.array(fit, pars = c("k_eco"))
sample_tZero <- as.array(fit, pars = c("tZero_eco"))
sample_phi <- as.array(fit, pars = "phi")

#Look at parameter space with intervals
# mcmc_intervals(sample_LInf)
# mcmc_intervals(sample_k)
# mcmc_intervals(sample_tZero)
# mcmc_intervals(sample_phi)

#Look at summary stats
# summary(fit, pars = c("LInf_eco"))
# summary(fit, pars = c("k_eco"))
# summary(fit, pars = c("tZero_eco"))
# summary(fit, pars = c("phi"))

#MCMC combo plots
png(filename = "Figures/Appendix/histtrace_LInf.png", width = 5000, height = 2500, res = 600)
mcmc_combo(samples, c("hist", "trace"), pars = c("LInf_eco[1]", "LInf_eco[2]", "LInf_eco[3]", "LInf_eco[4]"))
dev.off()

png(filename = "Figures/Appendix/histtrace_k.png", width = 5000, height = 2500, res = 600)
mcmc_combo(samples, c("hist", "trace"), pars = c("k_eco[1]", "k_eco[2]", "k_eco[3]", "k_eco[4]"))
dev.off()

png(filename = "Figures/Appendix/histtrace_tZero.png", width = 5000, height = 2500, res = 600)
mcmc_combo(samples, c("hist", "trace"), pars = c("tZero_eco[1]", "tZero_eco[2]", "tZero_eco[3]", "tZero_eco[4]"))
dev.off()

png(filename = "Figures/Appendix/histtrace_phi.png", width = 5000, height = 2500, res = 600)
mcmc_combo(samples, c("hist", "trace"), pars = c("phi"))
dev.off()

#Pairs plots
png(filename = "Figures/Appendix/pairs_eco1.png", width = 5000, height = 2500, res = 600)
pairs(fit, pars = c("LInf_eco[1]", "k_eco[1]", "tZero_eco[1]","phi"))
dev.off()

png(filename = "Figures/Appendix/pairs_eco2.png", width = 5000, height = 2500, res = 600)
pairs(fit, pars = c("LInf_eco[2]", "k_eco[2]", "tZero_eco[2]", "phi"))
dev.off()

png(filename = "Figures/Appendix/pairs_eco3.png", width = 5000, height = 2500, res = 600)
pairs(fit, pars = c("LInf_eco[3]", "k_eco[3]", "tZero_eco[3]", "phi"))
dev.off()

png(filename = "Figures/Appendix/pairs_eco4.png", width = 5000, height = 2500, res = 600)
pairs(fit, pars = c("LInf_eco[4]", "k_eco[4]", "tZero_eco[4]", "phi"))
dev.off()
```
